Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-07-29 paper/main.tex
remove todo
rewordings

\paragraph{Applications \& Parameter Estimation}
% \begin{itemize}
%     \item \cite{Winkle2017} Modeling mechanical interations in growing populations of rod-shaped
%         bacteria
%     \item \cite{Doumic2020} "A purely mechanical model with asymmetric features for early
%         morphogenesis of rod-shaped bacteria micro-colony"
%     \begin{itemize}
%         \item Really good paper for referencing
%         \item comparison of distributions; not individual agents
%         \item no bending in model
%         \item uses steric force (see also \cite{Trejo2013})
%     \end{itemize}
%     \item \cite{Grant2014} purposely-built model written in C++
%     \begin{itemize}
%         \item describes bacteria as collection of overlapping spheres
%         \item spheres are coupled by non-linear springs (Euler-Bernoulli dynamic beam theory). This
%             assumption is werid but does not alter their results
%         \item Only model repulsive forces; no attraction, adhesion
%     \end{itemize}
%     \item \cite{Cho2007} only 2D model; no param estimation; based on work done in \cite{Jnsson2005}
%     \item \cite{Storck2014} no bending; 41 parameters for various cases; only 8 parameters taken
%         from literature values or quantified; generate growth rate from normal distribution at each
%         integration step; stochastically equivalient for collection of bacteria but unfounded for
%         individual bactera
%     \item \cite{You2018} TODO
% \end{itemize}

@@

\paragraph{The missing Link}
It is clear that researchers are interested in studying mechanical properties of rod-like bacteria
in order to better understand phenomena such as biofilm formation.
The major downside of existing computational models is that there are no methods to properly
quantify their parameters which diminishes their explanatory power.
By leveraging a simple assumption about the cellular representation, this work aims to solve this
particular problem.

@@

Visualization methods are an integral part of any modeling endevaour.
They allow researchers to obtain feedback quickly and improve iteration times when constructing new
models.
Furthermore, we utilize them in our more advanced parameter estimation methods (see
section~\ref{subsec:comparison-across-generations}) where they enable us to effectively compare
two bacterial configurations across division events.
In the future, we are planning to build on the current state and generate realistic microscopic
images and time-series.
Together with the parameter estimation methods, they can then be used to provide large amounts of
mechanistic data for the training cell-segmentation and cell-tracking machine-learning algorithms.

>>>>

\paragraph{Applications \& Parameter Estimation}
% \begin{itemize}
%     \item \cite{Winkle2017} Modeling mechanical interations in growing populations of rod-shaped
%         bacteria
%     \item \cite{Doumic2020} "A purely mechanical model with asymmetric features for early
%         morphogenesis of rod-shaped bacteria micro-colony"
%     \begin{itemize}
%         \item Really good paper for referencing
%         \item comparison of distributions; not individual agents
%         \item no bending in model
%         \item uses steric force (see also \cite{Trejo2013})
%     \end{itemize}
%     \item \cite{Grant2014} purposely-built model written in C++
%     \begin{itemize}
%         \item describes bacteria as collection of overlapping spheres
%         \item spheres are coupled by non-linear springs (Euler-Bernoulli dynamic beam theory). This
%             assumption is werid but does not alter their results
%         \item Only model repulsive forces; no attraction, adhesion
%     \end{itemize}
%     \item \cite{Cho2007} only 2D model; no param estimation; based on work done in \cite{Jnsson2005}
%     \item \cite{Storck2014} no bending; 41 parameters for various cases; only 8 parameters taken
%         from literature values or quantified; generate growth rate from normal distribution at each
%         integration step; stochastically equivalient for collection of bacteria but unfounded for
%         individual bactera
%     \item \cite{You2018}
% \end{itemize}

@@

\paragraph{The missing Link}
It is clear that researchers are interested in studying mechanical properties of rod-like bacteria
in order to better understand phenomena such as biofilm formation.
The major downside of existing computational models is that there are almost no methods to properly
quantify their parameters which diminishes their explanatory power.
By leveraging a simple assumption about the cellular representation, this work aims to solve this
particular problem.

@@

Visualization methods are an integral part of any modeling endevaour.
They allow researchers to obtain feedback quickly and improve iteration times when constructing new
models.
Furthermore, we utilize them in our more advanced parameter estimation methods (see
section~\ref{section:parameter-estimation}) where they enable us to effectively compare
two bacterial configurations across division events.
In the future, we are planning to build on the current state and generate realistic microscopic
images and time-series.
Together with the parameter estimation methods, they can then be used to provide large amounts of
mechanistic data for the training cell-segmentation and cell-tracking machine-learning algorithms.
